---
title: "Course Project for STAT 8051"
author: "Charlie, Audrey, Lukas, Li"
date: "12/11/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pull in data

```{r}
test <- read.csv("test_2021.csv")
train <- read.csv("train_2021.csv")
```


## Library the packages
```{r}
library(lubridate)
library(glmm)
library(MASS)
library(caret)
library(ROSE)
library(cvTools)
library(ROCR)
```


## Diagnostics
```{r}
str(train)
```


The data is havily skewed
```{r}
table(train$fraud)
```

## Data processing 

### Creat year and month var using claim_date
```{r}
train$year <- format(parse_date_time(train$claim_date, orders = c("ymd", "mdy", "dmy")),format="%Y")
train$month <- months(as.Date(parse_date_time(train$claim_date, orders = c("ymd", "mdy", "dmy"))))
train$year <- as.factor(train$year)
train$month <- as.factor(train$month)
```

### Remain the first three digits of zip_code and treat it as factor variable
```{r}
train$zip_code <- floor(train$zip_code/100)
train$zip_code <- as.factor(train$zip_code)
```

### Factorization-characters and binaries to factors for modeling
```{r}
train$gender <- as.factor(train$gender)
train$marital_status <- as.factor(train$marital_status)
train$high_education_ind <- as.factor(train$high_education_ind)
train$address_change_ind <- as.factor(train$address_change_ind)
train$living_status <- as.factor(train$living_status)
train$claim_day_of_week <- as.factor(train$claim_day_of_week)
train$accident_site <- as.factor(train$accident_site)
train$witness_present_ind <- as.factor(train$witness_present_ind)
train$channel <- as.factor(train$channel)
train$policy_report_filed_ind <- as.factor(train$policy_report_filed_ind)
train$vehicle_category <- as.factor(train$vehicle_category)
train$vehicle_color <- as.factor(train$vehicle_color)
str(train)
```

### Dealing with NA in training data
```{r}
train <- na.omit(train)
```


### Dealing with outliers in trainning data

#### Age of the driver
```{r}
par(mfrow=c(1,2))
boxplot(train$age_of_driver)
boxplot(train$age_of_driver)
```


```{r}
age_threshold <- quantile(train$age_of_driver,0.99)
```

```{r}
tab.age <- table(train$fraud,train$age_of_driver>age_threshold)
prop.table(tab.age)
```

```{r}
chisq.test(tab.age,simulate.p.value = FALSE)
```

#### Annual income

```{r}
par(mfrow=c(1,2))
boxplot(train$annual_income)
boxplot(test$annual_income)
```


#### claim_est_payout

```{r}
par(mfrow=c(2,2))
boxplot(train$claim_est_payout)
boxplot(test$claim_est_payout)
```


## Data graphing

```{r}
plot(as.factor(train$fraud)~train$zip_code)
```

```{r}
tab.zip <- table(as.factor(train$fraud),train$zip_code)
chisq.test(tab.zip,simulate.p.value = TRUE)
```


## Logistic regression
### Stepwise AIC to select the best variable
```{r}
BigMod <- glm(fraud~age_of_driver+gender+
    marital_status+safty_rating+annual_income+high_education_ind+address_change_ind+
    living_status+accident_site+past_num_of_claims+witness_present_ind+
    liab_prct+channel+policy_report_filed_ind+claim_est_payout+age_of_vehicle+
    vehicle_price+vehicle_weight+year+month+zip_code,family=binomial,data=train)

stepAIC(BigMod, scope = list(upper=~age_of_driver+gender+
    marital_status+safty_rating+annual_income+high_education_ind+address_change_ind+
    living_status+accident_site+past_num_of_claims+witness_present_ind+
    liab_prct+channel+policy_report_filed_ind+claim_est_payout+age_of_vehicle+
    vehicle_price+vehicle_weight+year+month+zip_code,
    lower=~1),direction="both",trace=0,k=2,data=train)
```
### Cross-validation to choose cutoff point 
```{r}
dat <- train
n.fold <- 5
folds <- cvFolds(nrow(dat),n.fold)
result <- vector("list",n.fold)
for (k in 1:n.fold){
    dat.train <- dat[folds$subsets[folds$which!=k],]
    dat.test <- dat[folds$subsets[folds$which==k],]
    dat.train.over <- ovun.sample(fraud~.,data=dat.train,method = "over",p=0.5)$data
    mod <- glm(formula = fraud ~ age_of_driver+marital_status + gender + 
      safty_rating + annual_income + high_education_ind + address_change_ind + 
      living_status + accident_site + past_num_of_claims + witness_present_ind + 
      channel + claim_est_payout + age_of_vehicle + year+ zip_code, family =
      binomial, data = dat.train.over)
    predict.mod <- predict(mod,newdata = dat.test,type = "response")
    pred <- prediction(predict.mod,dat.test$fraud)
    per <- performance(pred,"f")
    result[[k]] <- per@x.values[[1]][which.max(per@y.values[[1]])]
}
```

```{r}
cutoffs.pre <- result
```

```{r}
cv.cutoff <- function(dat,n.fold,cutoff.list){
  folds <- cvFolds(nrow(dat),n.fold)
  p <- length(cutoff.list)
  result <- matrix(0,length(cutoff.list),n.fold)
  for (k in 1:n.fold){
    dat.train <- dat[folds$subsets[folds$which!=k],]
    dat.test <- dat[folds$subsets[folds$which==k],]
    dat.train.over <- ovun.sample(fraud~.,data=dat.train,method = "over",p=0.4)$data
    mod <- glm(formula = fraud ~ age_of_driver+marital_status + gender + 
      safty_rating + annual_income + high_education_ind + address_change_ind + 
      living_status + accident_site + past_num_of_claims + witness_present_ind + 
      channel + claim_est_payout + age_of_vehicle + year+ zip_code, family =
      binomial, data = dat.train.over)
    predict.mod <- predict(mod,newdata = dat.test,type = "response")
    for (j in 1:p){
      pred <- ifelse(predict.mod > cutoff.list[j],1,0)
      expected_value <- factor(dat.test$fraud)
      predicted_value <- factor(pred)
      CM <- confusionMatrix(data=predicted_value, reference = expected_value,positive = "1")
      re = CM$byClass[1]
      prec = CM$byClass[5]
      result[j,k] = 2 * prec * re / (prec + re)
    }
  }
  Fs <- apply(result,1,mean)
  return(list(cutoff.list = cutoff.list,F = Fs))
}
```


```{r}
cutoffs.op <- numeric(50)
best_values <- numeric(50)
for (k in 1: 50){
  cutoff.list <- seq(0.39,0.41,0.01)
  cutoffs <- cv.cutoff(train,5,cutoff.list)
  cutoffs.op[k] <- cutoffs$cutoff.list[which.max(cutoffs$F)]
  best_values[k] <- max(cutoffs$F)
}
```

```{r}
hist(cutoffs.op)
```

```{r}
cutoff.list <- seq(0.25,0.35,0.01)
cutoffs <- cv.cutoff(train,5,cutoff.list)
cutoff.op <- cutoffs$cutoff.list[which.max(cutoffs$F)]
```


### Fitting to test data
```{r}
#Organize test in the same way as train
test$gender <- as.factor(test$gender)
test$marital_status <- as.factor(test$marital_status)
test$high_education_ind <- as.factor(test$high_education_ind)
test$address_change_ind <- as.factor(test$address_change_ind)
test$living_status <- as.factor(test$living_status)
test$zip_code <- floor(test$zip_code/100)
test$zip_code <- as.factor(test$zip_code)
test$claim_day_of_week <- as.factor(test$claim_day_of_week)
test$accident_site <- as.factor(test$accident_site)
test$witness_present_ind <- as.factor(test$witness_present_ind)
test$channel <- as.factor(test$channel)
test$policy_report_filed_ind <- as.factor(test$policy_report_filed_ind)
test$vehicle_category <- as.factor(test$vehicle_category)
test$vehicle_color <- as.factor(test$vehicle_color)
test$year <- format(parse_date_time(test$claim_date, orders = c("ymd", "mdy", "dmy")),format="%Y")
test$month <- months(as.Date(parse_date_time(test$claim_date, orders = c("ymd", "mdy", "dmy"))))
test$year <- as.factor(test$year)
test$month <- as.factor(test$month)
test$marital_status[is.na(test$marital_status)] <- 0
test$witness_present_ind[is.na(test$witness_present_ind)] <- 0
test$claim_est_payout[is.na(test$claim_est_payout)] <- mean(na.omit(test$claim_est_payout))
test$age_of_vehicle[is.na(test$age_of_vehicle)] <- mean(na.omit(test$age_of_vehicle))
```


```{r}
train.over <- ovun.sample(fraud~.,data=train,method = "over",p=0.4)$data
```

```{r}
mod <- glm(formula = fraud ~ age_of_driver+marital_status + gender + 
      safty_rating + annual_income + high_education_ind + address_change_ind + 
      living_status + accident_site + past_num_of_claims + witness_present_ind + 
      channel + claim_est_payout + age_of_vehicle + year+ zip_code, family =
      binomial, data = train.over)
newval <- predict(mod,newdata=test,type="response")
submit <- ifelse(newval>0.4,1,0)
mat <- cbind(claim_number=test$claim_number,pred=submit)
```

Writing into CSV format for submission
```{r}
write.csv(mat,file="Prediction - Logistic Regression.csv",row.names = FALSE)
```
